Classless Java: Tuning Java Interfaces

### 2016-03-10 OOPSLA16/sections/Introduction.tex
Some edits and comments

Object oriented languages always strive to offer great code reuse,
they aim to couple flexibility and rigour, expressive power and modular reasoning.
Two main idea have emerged to this end: prototype based (PB) and class based (CB) languages.
In prototype based languages an objects inherits from other objects,
and thus objects own both behaviour and state (and objects is all you have).
In class based languages an object is instance of a specific class, and classes inherits from other classes.
Here objects own the state, while classes contains behaviour and the structure of the state.

>>>>

\bruno{Generaly speaking I like the flow of the new introduction!
I would say that we do need to defend ourselfves against Scala traits, 
but the differences are probably in three aspects:

\begin{itemize}

\item The abstract state operations and the novel technique to deal with 
covariant refinement of setters; Scala traits do support state (variables and 
immutable values), but there's no support for ``with'' (and clone-like) methods for example; and 
mutable state does not support type-refinement.

\item Automatic type-refinements (although Scala does also support \emph{manual} type-refinement 
of immutable fields);

\item We support constructors; whereas Scala traits do not. This is one key reason why classes 
are still necessary in a lnaguage like Scala.

One final point about the introduction is that language tuning is no longer 
mentioned. I think this is fine, but we have to discuss the strategy.

}
Object oriented languages always strive to offer great code reuse,
they aim to couple flexibility and rigour, expressive power and modular reasoning.
Two main idea have emerged to this end: prototype based (PB) and class based (CB) languages.
In prototype based languages an objects inherits from other objects,
and thus objects own both behaviour and state (and objects is all you have).
In class based languages an object is instance of a specific class, and classes inherits from other classes.
Here objects own the state, while classes contains behaviour and the structure of the state.

### 2016-03-10 OOPSLA16/sections/Overview.tex
Some edits and comments

This section illustrates how our programming style, supported by the \mixinAnn{} annotation, enable powerful
idioms similar to multiple inheritance and field type refinement.
We propose a standard example: \Q@Animal@s with a two-dimensional \Q@Point2D@ representing their \Q@location@.
Some kinds of animals are \Q@Horse@s and \Q@Bird@s.
Birds can \Q@fly@, thus their locations need to be three-dimensional \Q@Point3D@s (field type refinement).
Finally, we model \Q@Pegasus@ (one of the best-known creatures in Greek
mythology) as a kind of \Q@Animal@ with the skills
of both \Q@Horse@s and \Q@Bird@s (multiple inheritance). A simple class diagram illustrating
the basic system is given on the left side of Figure~\ref{fig:pegasus}.\footnote{Some research argues in favor of using subtyping for modeling taxonomies, other research argues against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}

@@

\paragraph{\Q@Point2D@: simple immutable data with fields}
Here we will illustrate how points are modelled with
interfaces. In IB we do not talk about state directly; state is accessed
using abstract (getter) methods. Note how in   The usual
approach to model points in Java would be to use a class with fields for the
coordinates.

@@

\paragraph{\Q@Point3D@ and properties updaters}
The \Q@Point3D@ interface is defined as follows:

@@

\subsection{Advanced Multiple Inheritance with Object Interfaces}
Finally, defining \texttt{Pegasus} is as simple as we did in the simplified
(and stateless) version on the right of Figure~\ref{fig:pegasus}.
 Note how even the non-trivial pattern for field type refinement is
transparently composed, and \texttt{Pegasus} has a \Q@Point3D@
\Q@location@.%  This works because \Q@Horse@ do not perform
% any field type refinement, otherwise we may have to choose/create a
% common subtype in order for \texttt{Pegasus} to exists.

@@

\subsection{Abstract state operations}
Defining an convenient list of abstract state operation is not an easy task.
Of course we need getters.
Currently we are offering both setters and Fluent setters. May be fluent setters are enough, and we can drop the
concept of setters?
We offer with withers and the with.
What can be a nice syntax?
\marco{my coments from now on}
A language natively supporting Object interfaces may offer the following syntax:
\begin{lstlisting}
obj interface Foo{
  int bar;//just getter
  var int bar;//also setter(fluent)
  with int bar;//also wither
  with var int bar;//getter setter wither
  with/var int bar=2;//default value, not asked in of
  invariant{/*stuff that should always hold*/}
  updater Bar, Beer//multi setter
  wither Foo, Beer//multi with
  factory of(...){call generate of}// generated of is now private
}
\end{lstlisting}

>>>>

This section illustrates how our programming style, supported by the \mixinAnn{} annotation, enable powerful
idioms similar to multiple inheritance and field type refinement.
We propose a standard example: \Q@Animal@s with a two-dimensional \Q@Point2D@ representing their \Q@location@.
Some kinds of animals are \Q@Horses@ and \Q@Bird@s.
Birds can \Q@fly@, thus their locations need to be three-dimensional \Q@Point3D@s (field type refinement).
Finally, we model \Q@Pegasus@ (one of the best-known creatures in Greek
mythology) as a kind of \Q@Animal@ with the skills
of both \Q@Horse@s and \Q@Bird@s (multiple inheritance). A simple class diagram illustrating
the basic system is given on the left side of Figure~\ref{fig:pegasus}.\footnote{Some research argues in favor of using subtyping for modeling taxonomies, other research argues against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
\bruno{Should we provide a one sentence summary in the abstract of how much code 
is needed in Java (without Obj) vs the approach with CJ, for the Pegasus example?}

@@

\paragraph{\Q@Point2D@: simple immutable data with fields}
Here we will illustrate how points are modelled with
interfaces. In IB we do not talk about state directly; state is accessed
using abstract (getter) methods. \bruno{unfinished sentence: Note how in}   The usual
approach to model points in Java is to use a class with fields for the
coordinates.

@@

\bruno{At some point we should talk about the point that was discussed in the reviews}
\paragraph{\Q@Point3D@ and properties updaters}
The \Q@Point3D@ interface is defined as follows:

@@

\subsection{Advanced Multiple Inheritance}
Finally, defining \texttt{Pegasus} is as simple as we did in the simplified
(and stateless) version on the right of Figure~\ref{fig:pegasus}.
 Note how even the non-trivial pattern for field type refinement is
transparently composed, and \texttt{Pegasus} has a \Q@Point3D@
\Q@location@.%  This works because \Q@Horse@ do not perform
% any field type refinement, otherwise we may have to choose/create a
% common subtype in order for \texttt{Pegasus} to exists.

@@

\subsection{Abstract state operations}\bruno{Is this something to go into a discussion later on?}
Defining an convenient list of abstract state operation is not an easy task.
Of course we need getters.
Currently we are offering both setters and Fluent setters. May be fluent setters are enough, and we can drop the
concept of setters?
We offer with withers and the with.
What can be a nice syntax?
\marco{my coments from now on}
A language natively supporting Object interfaces may offer the following syntax:
\begin{lstlisting}
obj interface Foo{
  int bar;//just getter
  var int bar;//also setter(fluent)
  with int bar;//also wither
  with var int bar;//getter setter wither
  with/var int bar=2;//default value, not asked in of
  invariant{/*stuff that should always hold*/}
  updater Bar, Beer//multi setter
  wither Foo, Beer//multi with
  factory of(...){call generate of}// generated of is now private
}
\end{lstlisting}
